Hopf Algebras\vspace{-1ex}

### 2019-02-13 Hopf Algebra Notes/Hopf Algebras.tex
I should push more frequently

\DeclareMathOperator{\Ext}{Ext}

@@

%Hopf algebra stuff
\newcommand*{\Vectk}{\operatorname{Vect}_k}
\newcommand*{\Algk}{\operatorname{Alg}_k}
\newcommand*{\Coalgk}{\operatorname{Coalg}_k}
\DeclareMathOperator{\Coalg}{Coalg}
\DeclareMathOperator{\Alg}{Alg}
\newcommand*{\scrT}{\ensuremath\mathscr{T}}

@@

\begin{ex}
	Consider $V\in _V^V\mathcal{V}$. THen you can define $\rho=\Delta$ and check it satisfies all the requirements.
\end{ex}
\begin{defn}
	An $\binom{l}{l}$ Hopf $V$-module $M$ is called \textbf{trivial} if $M\cong V\otimes M_0$ for some $M_0\in\Vectk.$
\end{defn}
\begin{rmk}
	Basically you can think of this being trivial since we can \textit{always} define a module in this way
	where the entire module structure is inherited from the structure of $V$ (that is, irrespective of $M_0$).
\end{rmk}
\begin{thm-prime}{thm-LS69}%[Fundamental Thm of Hopf Modules (L-S `69)]
	Suppose that $V$ is a Hopf algebra. Then
	\begin{enumerate}
		\item Every $\binom{l}{l}$ Hopf $V$-module is trivial.
		\item Let $M\in _V^V\mathcal{M}$. Then 
		\[M\xrightarrow{\sim} V\otimes M^{Cov}\]
		where
		\[M^{Cov}:=\{m\in M|\rho(m)=1\otimes m\}.\]
		\item $_V^V\mathcal{M}\cong \Vectk$.
	\end{enumerate}
\end{thm-prime}
\begin{rmk}
	This is actually just a reformulation of theorem~\ref{thm-LS69} in less compact (but ultimately more usable) notation.
\end{rmk}
\begin{prf}
	Some identites that will be useful:
	\begin{align*}
		\Delta(v)=\sum v_1\otimes v_2\\
		\sum S(v_1)\otimes v_2=\varepsilon(v)=\sum v_1S(v_2)\\
		\rho(m)=\sum m_{-1}\otimes m_1\\
		\sum v_1\otimes (v_2)_1\otimes (v_2)_2=\sum (v_1)_1\otimes (v_1)_2\otimes v_2=\sum v_1\otimes v_2\otimes v_3\\
		\sum (m_{-1})_1\otimes (m_{-1})_2\otimes m_0=\sum m_{-1}\otimes (m_0)_{-1}\otimes (m_0)_0=\sum m_{-2}\otimes m_{-1}\otimes m_0
	\end{align*}

@@

	Finally the last claim will be seen on Friday.
\end{prf}
\end{document}

>>>>

\DeclareMathOperator{\Ext}{Ext}
\DeclareMathOperator{\gldim}{gldim}
\DeclareMathOperator{\projdim}{projdim}
\DeclareMathOperator{\findim}{findim}

@@

%Hopf algebra stuff
\newcommand*{\Vectk}{\operatorname{Vect}_k}
\newcommand*{\Algk}{\operatorname{Alg}_k}
\newcommand*{\Coalgk}{\operatorname{Coalg}_k}
\newcommand*{\calM}{\mathcal{M}}
\DeclareMathOperator{\Coalg}{Coalg}
\DeclareMathOperator{\Alg}{Alg}
\newcommand*{\scrT}{\ensuremath\mathscr{T}}
\newcommand*{\Ch}{\mathbf{Ch}}

@@

\begin{ex}
	Consider $V\in _V^V\mathcal{M}$. THen you can define $\rho=\Delta$ and check it satisfies all the requirements.
\end{ex}
\begin{defn}
	An $\binom{l}{l}$ Hopf $V$-module $M$ is called \textbf{trivial} if $M\cong V\otimes M_0$ for some $M_0\in\Vectk.$
\end{defn}
\begin{rmk}
	Basically you can think of this being trivial since we can \textit{always} define a module in this way
	where the entire module structure is inherited from the structure of $V$ (that is, irrespective of $M_0$).
\end{rmk}
\begin{thm-prime}{thm-LS69}%[Fundamental Thm of Hopf Modules (L-S `69)]
	Suppose that $V$ is a Hopf algebra. Then
	\begin{enumerate}
		\item Every $\binom{l}{l}$ Hopf $V$-module is trivial.
		\item Let $M\in _V^V\mathcal{M}$. Then 
		\[M\xrightarrow{\sim} V\otimes M^{Cov}\]
		where
		\[M^{Cov}:=\{m\in M|\rho(m)=1\otimes m\}.\]
		\item $_V^V\mathcal{M}\cong \Vectk$.
	\end{enumerate}
\end{thm-prime}
\begin{rmk}
	This is actually just a reformulation of theorem~\ref{thm-LS69} in less compact (but ultimately more usable) notation.
\end{rmk}
\begin{prf}
	Some identites that will be useful:
	\begin{align*}
		\Delta(v)=\sum v_1\otimes v_2\\
		\sum S(v_1)\otimes v_2=\varepsilon(v)=\sum v_1S(v_2)\\
		\rho(m)=\sum m_{-1}\otimes m_1\\
		\sum v_1\otimes (v_2)_1\otimes (v_2)_2=\sum (v_1)_1\otimes (v_1)_2\otimes v_2=\sum v_1\otimes v_2\otimes v_3\\
		\sum (m_{-1})_1\otimes (m_{-1})_2\otimes m_0=\sum m_{-1}\otimes (m_0)_{-1}\otimes (m_0)_0=\sum m_{-2}\otimes m_{-1}\otimes m_0
	\end{align*}

@@

	Finally the last claim will be seen on Friday.
\end{prf}
\section{January 18, 2019}
Recall the Fundamental Theorem of Hopf Modules: $_V^V\mathcal{M}\simeq \Vectk$ or equivalently that
every (not necessarily finite dimensional) Hopf module is trivial: $M\cong V\otimes M^{Cov}$.

\section{Frobenius Algebras}
Today we will discuss the result that
\begin{thm}[Larson-Sweedler, `69]\label{LS-69-2}
	Every finite dimensional Hopf algebra is Frobenius.
\end{thm}

There are some nice categorical equivalences: finite dimensional Frobenius algebras over $k$ is
equivalent to 2-D quantum field theories, or equivalently the symmetric functor category from the 2-D cobordism category to $\Vectk$.

Whoa.

\begin{ex}
	$V=(kG)^*=\oplus_{g\in G}k\delta g$ where $G$ is a finite group. Define the algebra structure with
	\[1_V=\sum_{g\in G}\delta_g\]
	and $\delta_g\delta_h=\delta_g$ when $g=h$ and 0 otherwise.

	The coalgebra structure is $\varepsilon(\delta_g)=\delta_{g,1_G}$ where on the right it is the Kroneker delta.
	Furthermore $\Delta(\delta_g)=\sum_{h\in G}\delta_h\otimes\delta_{h^{-1}g}$ and $S(\delta_g)=\delta_{g^{-1}}$.
\end{ex}

\begin{prob}
	Prove that $kS_3\cong(kS_3)^*$.
\end{prob}
\begin{prob}
	Prove that $k\Z_3\cong(k\Z_3)^*$ when $\ch k\ne 3$.
\end{prob}

\begin{lem}
	Let $A$ and $B$ be algebras over $k$.
	\begin{enumerate}
		\item If $M\in_A\mathcal{M}_B$, $N\in \mathcal{M}_B$, then $\Hom_{\mathcal{M}_B}(_AM,N)\in\mathcal{M}_A$.
		\item Take $B=k$, $M\in _A\mathcal{M}$. Then
		\[M^*:=\Hom_k(M,k)\in\mathcal{M}_A.\]
		\item $(-)^*:_A\mathcal{M}_{f.d.}\to (\mathcal{M}_A)_{f.d.}^{op}$ is an equivalence of categories. This is called the \textbf{reflection principle.}
	\end{enumerate}
\end{lem}
\begin{prf}
	This is a known fact. (b) follows quickly from (a) and (c) involves extending the dual map to maps of left $A$-modules.
\end{prf}
By the principle of duality, we have:
\begin{lem}
	If $C$ and $d$ are finite dimensional coalgebras over $k$, then
	\begin{enumerate}
		\item if $M\in ^C\mathcal{M}^D_{f.d.}$ and $N\in \mathcal{M}^D_{f.d.}$ then $\Hom_{\mathcal{M}^D}(M,N)\in\mathcal{M}^C_{f.d.}$.
		\item If $M\in ^C\mathcal{M}_{f.d.}$ then $\Hom_k(M,k)\in\mathcal{M}^C$
		\item $(-)^*:^C\mathcal{M}_{f.d.}\to(\mathcal{M}_{f.d.}^C)^{op}$ is an equivalence of categories.
	\end{enumerate}
\end{lem}

\begin{defn}
	A finite dimensional algebra $A$ is called \textbf{Frobenius} if one of the following (equivalent)
	conditions hold:
	\begin{enumerate}
		\item $(A)^*_A\cong A_A$
		\item $_A(A)^*\cong _AA$
	\end{enumerate}
\end{defn}
\begin{rmk}
	Note that the equivalence of these two things follow due to the equivalence of categories we get
	from the two lemmas above.
\end{rmk}
We actually don't need the following lemma, but James is a fan so we're going to write it down.
\begin{lem}\label{lem-not-needed}
	Let $A$ be a finite dimensional algebra over $k$. Then $M$ is a finite dimensional left $A$-module if
	and only if $M$ is a finite dimensional right $A^*$-comodule.
\end{lem}
\begin{rmk}
	The reason we get the module to comodule switch is (partly?) due to the fact that in our
	monoidal category $(V\otimes W)^*=W^*\otimes V^*.$
\end{rmk}

\begin{lem}
	Let $V$ be a finite dimensional Hopf algebra and $M$ be a finite dimensional $\binom{l}{l}$ Hopf $V$-module.
	Then $M^*$ is a finite dimensional $\binom{r}{r}$ Hopf $V$-module.
\end{lem}
\begin{prf}
	$M\in {_V^V\mathcal{M}}$ if and only if $M\in_V\mathcal{M}\cap {^V\mathcal{M}}$ and the two structures are compatible.

	The fact that $M^*\in \mathcal{M}_V\cap\mathcal{M}^V$ is straightforward and checking the compatibility is easy to check.
\end{prf}

Now we finally prove that theorem~\ref{LS-69-2}:
\begin{prf}
	First of all $V\in{^V_V\mathcal{M}}$ so by the lemma $V^*\in\mathcal{M}_V^V$. By the fundamental theorem of Hopf modules,
	\[V^*\cong V\otimes(V^*)^{Cov}\cong V\]
	as Hopf modules. Thus $V_V^*\cong V_V$ in particular.
\end{prf}
\begin{rmk}
	The second isomorphism above comes from the fact that $(V^*)^{Cov}$ is one-dimensional, so is the trivial module. :)
\end{rmk}
\begin{cor}
	If $A$ is \textit{not} a Forbenius algebra, then there is \textit{no} Hopf algebra structure on $A$.
\end{cor}
\begin{ex}
	Some examples of algebras with(out) this property.
	\begin{enumerate}
		\item $\C[x]/(x^n)$ is finite dimensional and Frobenius, but if $n\ge 2$, there is \textbf{no Hopf structure.}
		\item $\C[x,y]/(x^2,y^2,xy)=\C1+\C x+\C y$. This has the property that it is local (unique maximal submodule over itself)
		but $A^*$ has two maximal submodules, so $A\not\cong A^*$, so not Frobenius.
	\end{enumerate}
\end{ex}
Finally we conclude with a proof of lemma~\ref{lem-not-needed}:
\begin{prf}
	Say that $\{a_i\}$ is a basis for $A$ and $\{a_i^*\}$ is a basis for $A^*$ ($a_i^*(a_j)=\delta_{ij}$).

	Define $\rho(m)=\sum(a_i\cdot m)\otimes a_i^*$. A lemma (use change of basis matrices) shows that
	$\sum a_i\otimes a_i^*$ is independent of choice of basis. Another lemma says (scalar) multiplication is associative
	if and only if $\rho$ is coassociative. This gets us the forward direction.

	For the reverse direction, say $\rho(m)=\sum m_0\otimes m_{-1}.$ then define $a\cdot m=\sum m_1(a)m_0$.
\end{prf}

\section{January 23, 2019}
Today we are interested in studying the representation category $_V\mathcal{M}$ for any Hopf algebra $V$. It will end up
(and this is a bit cryptic for now) that $k$ controls everything here.
\section{Represetnations and Modules}
Notice that for any Hopf algebra $V$ we have the exact sequence:
\[0\to \ker\varepsilon\to V\xrightarrow{\varepsilon} k\to 0\]
\begin{defn}
	The trivial $V$ modules is $V/\ker\varepsilon=k\in _V\mathcal{M}_V$.
\end{defn}
\begin{ex}
	$V=kG$ where $G$ is a group, $\varepsilon(g)=1$ for all $g\in G$. Then $\ker\varepsilon=\oplus k(g-1)$ and so
	\[k=V/\oplus k(g-1).\]
\end{ex}
\begin{ex}
	$V=k[x]$ where $\Delta(x)=1\otimes x+x\otimes 1$ and $\varepsilon(x)=0$. Then $\ker\varepsilon=\oplus_{n\ge 1}kx^n$, whence
	\[V/\ker\varepsilon=k[x]/(x).\]
\end{ex}
\begin{ex}
	Let $V=(kG)^*=\oplus_G k\delta_g$ and $\varepsilon(\delta_g)=\delta_{1,1_G}$. THen $\ker\varepsilon=\oplus_{g\ne 1_G}k\delta_g$ and
	\[V/\ker\varepsilon=k\delta_1\]
\end{ex}
The following uses $T_4$, the example of the day I missed due to my meeting.
\begin{ex}
	$T_4$: $\varepsilon(g)=1$, $\varepsilon(p)=0$. Then $\ker\varepsilon=k(g-1)\oplus kp\oplus kpg$ and
	\[V/\ker\varepsilon=k.\]
\end{ex}
\begin{lem}
	\begin{itemize}
		\item For all $x \in V$ and $a\in k$ (the trivial module), $xa=\varepsilon(x)a$.
		\item $k$ is the identity object in $_V\mathcal{M}$.
		\item If $V$ is cocommutative, then $_V\mathcal{M}$ is a symmetric tensor category.
	\end{itemize}
\end{lem}
\begin{prf}
	\[x\cdot a=(x-\varepsilon(x))a+\varepsilon(x)a=\varepsilon(x)a\]
	since $x-\varepsilon(x)\in\ker\varepsilon$.

	For the next part, consider that for every $M\in _V\mathcal{M}$, we have 
	\[M\otimes k\cong M\cong k\otimes M.\]
	But then for all $x\in V$, we have
	\begin{align*}
		\varphi[x\dot(m\otimes 1)]&=\varphi(x_1\cdot m\otimes x_2\cdot 1)\\
		&=\varphi(x_1\cdot m\otimes \varepsilon(x_2)\cdot 1)\\
		&=\varphi((x_1\varepsilon(x_2))m\otimes 1)\\
		&=\varphi(xm\otimes 1)=xm=x\varphi(m\otimes 1).
	\end{align*}
\end{prf}
\subsection{Integrals}
\begin{defn}
	An element $x\in V$ is called a \textbf{left integral} if $vx=\varepsilon(v)x$ for all $v\in V$.
	Similar for \textbf{right integral.}
\end{defn}
\begin{lem}
	The following are equivalent:
	\begin{itemize}
		\item $x\in V$ is a left integral.
		\item $kx\cong$ the trivial module.
		\item $1\mapsto x$ defines a left $V$-module morphism $k\to kx$.
	\end{itemize}
\end{lem}
\begin{lem}
	Let $\int_V^l$ denote the set of left integrals in $V$.
	\begin{itemize}
		\item $\int_V^l$ forms a vector subspace of $V$.
		\item $\int_V^l$ forms a left $V$ submodule of $V$. Thus it forms a left ideal of $V$.
		\item $\int_V^l$ forms a right ideal of $V$.
		\item $\int_V^l\cong\Hom_V(k,V)$ (left $V$-mods)
	\end{itemize}
\end{lem}
\begin{prf}
	The first three are relatively obvious. The last one comes from the map $\Phi:x\in\int_V^l\mapsto f_x:1\to x.$
\end{prf}
\begin{rmk}
	Unfortunately it ends up that the collection above ends up usually being zero. 
\end{rmk}
\begin{thm}
	Let $V$ be a finite dimensional Hopf algebra. Then $\dim\int_V^l=1$.
\end{thm}
\begin{prf}
	\begin{align*}
		\dim\int_V^l&= \dim\Hom_V(_Vk,_VV)\\
		&=\dim\Hom_V((V^*)_V,k_V^*)\\
		&=\dim\Hom_V(V_V,k)=\dim k=1
	\end{align*}
	Where we used above that $V$ was finite dimensional whence Frobenius to get $V^*\cong V$.
\end{prf}
\begin{ex}
	$V=kG$ for $G$ a finite group. Then $\int_V^l=\sum_G g$ and
	\[v\cdot(\sum_G g)=\varepsilon(v)(\sum_G g)=\int_V^r\]
	and when $v=h\in G$, we can compute it.
\end{ex}
\begin{ex}
	When $V=k[x]$, which is infinite dimensional, we get no integrals. This is because it is an 
	integral domain, so if
	\[x\cdot\int_V^l=\varepsilon(x)\int_V^l=0\cdot\int_V^l=0\]
	then this forces $\int_V^l=0$.
\end{ex}
\begin{ex}
	When $V=(kG)^*$ for $G$ a finite group, then 
	\[\int_V^l=\int_V^r=\delta_{1_G}\]
\end{ex}
\begin{ex}
	When $V=T_4=k1\oplus kg\oplus kp\oplus kgp$,
	\[\int_V^l=(g+1)p\ne\int_V^r=p(1+g)=(1-g)p\]
\end{ex}
\begin{ex}
	When $\ch k=p$ and $V=k[x]/(x^p)$, then 
	\[\int_V^l=\int_V^r=x^{p-1}.\]
\end{ex}

\section{January 28, 2019}
Today we will be mostly doing things in homological algebra. :)
\begin{ex}
	An example of an infinite dimension Hopf algebra is of the following: fix $q\in k^*$. Then 
	\[T_\infty=k<g, g^{-1}, p>/\langle gg^{-1}=g^{-1}g=1, gp=qpg\rangle\]
	and we can find $\{g^ip^j:i\in\Z, j\in\N\}$ is a $k$-linear basis.

	For the coalgebra structure, define $\Delta(g)=g\otimes g$ and $\varepsilon(g)=1$ and 
	$\Delta(p)=p\otimes 1+g\otimes p$ and $\varepsilon(p)=0$. Finally $S(g)=g^{-1}$ and $S(p)=q^{-1}pg^{-1}$.
\end{ex}

\subsection{Some nice results in Homological algebra}
\begin{prob}
	Prove that $\Q$ is an injective $\Z$-module.
\end{prob}
\begin{thm}[Auslander-Buchsbaum `59]
	Every local commutative algebra with finite global dimension is a UFD.
\end{thm}
Another result is about algebraic groups:
\begin{thm}
	If $\ch k=0$, then every Noetherian commutative Hopf algebra has finite global dimension
	and is a direct sum of integral domains, each of which is isomorphic.
\end{thm}
\begin{defn}[Semiprime]
	An ideal $I\lhd R$ is called \textbf{semiprime} if it is the intersection of (possibly infinitely many) primes.
\end{defn}
\begin{conj}
	Every Noetherian Hopf algebra of finite global dimension is semiprime (that is, 0 is semiprime in $V$).
\end{conj}

\begin{defn}[Connected Graded]
	An algebra $A$ is called \textbf{connected graded} if 
	\[A=k\oplus A_1\oplus A_2\oplus \cdots\]
	and $1_A=1_k\in k$, $A_iA_j\subseteq A_{i+j}$.

	The new part (connected) refers to the fact that $k$ is a summand.
\end{defn}
\begin{thm}
	Suppose that $\ch k=0$. Let $V$ be a Noetherian Hopf algebra that is connected graded as an algebra (says nothing about the coalgebra structure).
	Then the following hold:
	\begin{itemize}
		\item $V$ has finite global dimension.
		\item $V$ is a domain.
		\item $V$ is Artin-Schelter regular.
		\item $V$ is Auslander regular and Cohen-Macaulay.
		\item $V$ is Calabi-Yau
		\item $V$ is an iterated Ore extension.
	\end{itemize}
\end{thm}
That's a lot of word salad.

\subsection{Projective Modules}
\begin{defn}[Module-Theoretic Definitions]
	Let $M$ be a (left) $A$ module. 
	\begin{enumerate}
		\item $M$ is \textbf{free} if $M\cong \oplus_\mathcal{I} A$
		\item $M$ is \textbf{projective} if it is a direct summand of a free module.
	\end{enumerate}
\end{defn}
\begin{lem}
	$p\in{_A\mathscr{M}}$ is projective if and only if the diagram in figure~\ref{fig-proj2} commutes.
\end{lem}
\begin{figure}[h]
	\centering
	\begin{tikzcd}
		& P\ar[d,"f"]\ar[dl,dashed,"F",swap] &\\
		M\ar[r,"\pi"] & N\ar[r] & 0
	\end{tikzcd}
	\caption{The definition of a projective object.}
	\label{fig-proj2}
\end{figure}
\begin{rmk}
	This actually is just the definition for any Abelian category. This may generalize even further, but this is enough for us. :)
\end{rmk}

\begin{defn}[Injective Module]
	An \textbf{injective} $V$-module is a projective $V^{op}$ module.
\end{defn}

\section{January 30, 2019}
Today we're talking more about homological algebra. In particular, we'll learn about (or see again)
\begin{itemize}
	\item Complexes
	\item Projective resolutions and dimension
\end{itemize}
First the example of the day:
\begin{ex}
	Consider $GL_2(k)$, a group -- in fact, an algebraic group! Then 
	\[\mathcal{O}(GL_2)=k[x_{11},x_{12},x_{21},x_{22},det^{-1}]/\langle det^{-1}(x_{11}x_{22}-x_{21}x_{12}-1\rangle\]
	Let $X=(x_{ij})\in GL_2(k)$.

	Define the coalgebra structure via $\Delta(X)=X\otimes X$ and $\Delta(x_{ij})=\sum_1^2 x_{is}\otimes x_{sj}$,
	$\varepsilon(X)=I_2$, $\varepsilon(x_{ij})=\delta_{ij}$. Some computation (this is not trivial!) give us
	\[\Delta(det)=det\otimes det\quad\Rightarrow\quad \Delta(det^{-1})=det^{-1}\otimes det^{-1}\]
	Then we can find that $S(X)=X^{-1}$ and $S(det^{\pm})=det^{\mp}$.
\end{ex}
\begin{prob}
	Construct $\mathcal{O}(GL_n)$.
\end{prob}
\begin{prob}
	Construct $\mathcal{O}(G)$.
\end{prob}
\subsection{Homological stuff}
\begin{defn}[Complex]
	A \textbf{complex} of $A$-modules is a sequence of $A$-modules connected by homomorphisms
	\[\cdots\to M_2\xrightarrow{d_2}M_1\xrightarrow{d_1} M_0\to \cdots\]
	where $d_i\circ d_{i+1}=0$ for all $i$.
\end{defn}
Then we went over the standard definitions for:
\begin{itemize}
	\item Homology
	\item (Short) Exact sequences
	\item Projective resolutions
	\item Projective dimension
\end{itemize}
\begin{defn}[Global (Homological) Dimension]
	The (left) \textbf{global dimension} is
	\[\gldim A=\max\{\projdim M| M\in {_A\mathcal{M}}\}\]
\end{defn}
\begin{thm}[Hilbert Syzygy Theorem]
	\[\gldim k[x_1,\dots,x_n]=n\]
\end{thm}
\begin{cor}
	\[\gldim k= 0\]
\end{cor}
\begin{defn}
	The \textbf{finitistic global dimension} of $A$ is defined to be
	\[\findim A=\max\{\projdim M| M\in{_A\mathcal{M}}, M\text{ is f.g.},\projdim M <\infty\}\]
\end{defn}
\begin{conj}[Finitistic Dimension Conjecture]
	Let $A$ be a finitely generated algebra. Then $\findim A<\infty$.
\end{conj}
\section{February 1, 2019}
A really cool theorem:
\begin{thm}[Lorenz-Lorenz `95]
	Let $V$ be a Hopf algebra. Then $\gldim V=\projdim {_Vk}$.
\end{thm}
\begin{prf}
	It is enough to show that $\gldim V\le\projdim{_Vk}$. Or equivalently that
	\[\projdim M\le \projdim k,\forall M\in{_V\calM}\]
	but then
	\[\projdim(M)=\projdim(k\otimes M)\le \projdim(k)\]
	by lemma~\ref{lem-tensorproj}
\end{prf}
\begin{ex}
	Today's example: Quantum group $V=\mathcal{O}_q(GL_2)$. Fix some $q\in k^*$. Then as an algebra, 
	\[V=k\langle x_{11},x_{12},x_{21},x_{22},det_q^{-1}\rangle/(\text{relations})\]
	where the relations are given by 
	\begin{align*}
		x_{ij}x_{kl}=(\delta_{ik}+\delta_{jl})x_{kl}x_{ij}\quad\text{except when } i=k \text{ and } j=l\\
		x_{12}x_{21}=x_{21}x_{12}\\
		x_{22}x_{11}-x_{11}x_{22}=(q-q^{-1})X_{12}x_{21}\\
		det_q^{-1}det_q=det_q det_q^{-1}=1
	\end{align*}
	where $det_q=x_{22}x_{11}-q^{-1}x_{12}x_{21}x_{11}x_{22}-qx_{12}x_{21}$. One can show $det_q$ is central in $V$.

	Now for the coalgebra: $\delta(X)=X\otimes X$, so $\Delta(x_{ij})=\sum_{s=1}^2 x_{is}\otimes x_{sj}$, $\varepsilon(X)=I_2$ and
	\[S(X)=X^{-1}=\begin{pmatrix}
		x_{22}det_q^{-1} & -qx_{12}det_q^{-1}\\
		-q^{-1}x_{21}det_q^{-1} & x_{11}det_q^{-1}
	\end{pmatrix}\]
\end{ex}
\begin{prob}
	Show that $\gldim(\begin{smallmatrix}\Z & \Q\\0 & \Q\end{smallmatrix})=2$ while $\gldim(\begin{smallmatrix}\Z & \Q\\0 & \Q\end{smallmatrix})^{op}=1$
\end{prob}

\subsection{Returning to Homological Algebra}
Let $A$ be an algebra, $M\in {_A\mathcal{M}}$. Recall the definitions of projective and global dimension.
\begin{rmk}
	The following are equivalent:
	\begin{enumerate}
		\item $\gldim A=0$
		\item $\gldim A^{op}=0$
		\item $A$ is semisimple Artininan
		\item $A=\oplus_1^b M_{n_i}(D_i)$ by Artin-Wedderburn where $D_i$ is a division ring.
	\end{enumerate}
\end{rmk}
\begin{rmk}
	\begin{enumerate}
		\item A PID has global dimension 1. 
		\item The free algebra $k\langle x_1,\dots,x_n\rangle$ has global dimension 1.
		\item The path algebra of a finite quiver has global dimension 1.
		\item $\mathcal{O}_q(GL_2)$ has global dimension 4.
	\end{enumerate}
\end{rmk}

From now on, let $V$ always stand for some Hopf algebra.
\begin{lem}
	${_V\mathcal{M}}$ is a tensor category. The action on $M\otimes N$ is the following:
	\[v\cdot (m\otimes n)=\sum (v_{(1)}\cdot m)\otimes (v_{(2)}\cdot n)\]
\end{lem}
\begin{lem}
	If $M\in {_V\mathcal{M}}$, then $V\otimes M\in {_V^V\mathcal{M}}$. As a corollary,
	\[V\otimes M\cong V\otimes (V\otimes M)^{cov}\]
	and as a consequence $V\otimes M$ is a free $V$-module.
\end{lem}

\begin{thm}
	Let $P\in {_V\calM}$ be projective. Then $P\otimes N\in {_V\calM}$ is projective for all $M\in{_V\calM}$
\end{thm}
\begin{prf}
	To see this, notice that $P$ is a summand of $V^{\oplus b}$. But then if $P\oplus Q\cong V^{b}$, then 
	\[[P\otimes N]\oplus[Q\otimes N]=[P\oplus Q]\otimes N=(V^b)\otimes N=(V\otimes N)^b\]
	and by the last lemma, this is free.
\end{prf}

\begin{lem}\label{lem-tensorproj}
	\[\projdim(M\otimes N)\le \min\{\projdim M,\projdim N\}\]
\end{lem}
\begin{prf}
	We only show that $\projdim M\otimes N\le \projdim M$. If $\projdim M=\infty$, we are done.
	If it is finite, take any minimal projective resolution of $M$ and tensor with $N$. This is a projective 
	resolution of $M\otimes N$ of length at most $\projdim M$. 
\end{prf}
\begin{rmk}
	The fact it is still exact holds because we can appeal to the vector space structure.
\end{rmk}

Some more results (mostly for finite dimensional algebras):
\begin{thm}[Radford `75]
	If $V$ is finite dimensional, then the antipode $S$ has finite order:
	\[S^d=\id_V\]
\end{thm}
\begin{thm}[Larson-Radford `88]
	Suppose $\ch k\ne 0$ (can be relaxed but gets more ugly). Then the following are equivalent:
	\begin{enumerate}
		\item $\gldim V=0$ ($V$ is semisimple Artinian)
		\item $\gldim V^*=0$
		\item $S^2=\id_V$
	\end{enumerate}
\end{thm}
\begin{thm}[Nichols-Zoeller `89]
	Let $V$ be finite dimensional. If $Q$ is a Hopf subalgebra of $V$, then $_WV$ and $V_W$ are free.
\end{thm}

\section{February 6, 2019}
First some history: in 1899, Maschke proves the following:
\begin{thm}
	Let $G$ be a finite group. Then $kG$ is semisimple (Artinian) if and only if $\ch k\nmid |G|$.
\end{thm}
Today, we will see an analog of Maschke for Hopf algebras:
\begin{thm}[``Maschke's Theorem'' (Larson-Sweedler `69)]
	Let $V$ be a finite-dimensional Hopf algebra. Then the following are equivalent:
	\begin{enumerate}
		\item $\gldim V = 0$ ($V$ is semisimple Artinian)
		\item $\varepsilon(\int^l)\ne 0$
		\item $\varepsilon(\int^r)\ne 0$
	\end{enumerate}
\end{thm}

Now say $V=kG$. Then $\int=\sum_{g\in G} g$ is both a left and right integral of $V$. This uses that $\varepsilon(g)=1$.

But then if $\varepsilon(\int)=\sum_G \varepsilon(g)=\sum_G 1=|G|\ne 0\Leftrightarrow \ch k\nmid |G|$.

\subsection{Example of the Day}
\begin{ex}[Kac-Paljutkin Algebra]
	$k=\C$ or any field with $\ch k\ne 2$. Then $V=k\langle x,y,z\rangle$ modulo the relations
	\begin{align*}
		x^2=y^2=1\\
		xy=yx\\
		xz=yz\\
		xy=xz\\
		x^2=\frac{1}{2}(1+x+y-xy)
	\end{align*}
	Then $V$ has a $k$-basis $\{1,x,y,z,xy,xz,yz,xyz\}$, so as an algebra $V\cong k^4\oplus M_2(k)$.

	Then $\Delta(x)=x\otimes x$, $\Delta(y)=y\otimes y$, $\Delta(z)=(1\otimes 1+1\otimes x+y\otimes 1-y\otimes x)(z\otimes z)$.

	And $\varepsilon(x)=\varepsilon(y)=\varepsilon(z)=1$ and $S(x)=x$, $S(y)=y$ and $S(z)=z$. Recall that although $S$ looks like the identity map, 
	$S$ is an \textbf{antihomomophism}. So actually $S^2=\id_V$.

	Claim:
	\[\int=1+x+y+z+xy+xz+yz+xyz\]
	then $\varepsilon(\int)=8\ne 0$, so $V$ is semisimple.

	Some nice facts:
	\begin{enumerate}
		\item $V^*\cong V$
		\item $V$ is the unique eight dimensional noncommutative, noncocommutative semisimple Hopf algebra.
		\item Let $Q_8$ be the quaternion group. Then $kQ_8$ is eight dimensional. Therefore $V\cong kQ_8$ as algebras, but \textbf{not as coalgebras.}
		The reason behind this is that $kG$ is \textit{always} cocommutative.
	\end{enumerate}
\end{ex}

\subsection{HW of the day}
\begin{prob}
	If $V$ is eight dimensional and noncommutative, then $V\cong k^4\oplus M_2(k)$ as an algebra. Thus if $G$ is nonabelian group of order 8, then $\C G\cong \C^4\oplus M_2(\C)$.
\end{prob}

\subsection{Proof of ``Maschke's theorem''}
Now we actually do the proof:
\begin{prf}
	(a) $\Rightarrow$ (b): Say the global dimension of $V$ is zero. Then $V$ is semisimple by Artin-Wedderburn and we have a decompositon
	\[V\cong \oplus_1^s M_{n_i}(D_i)\]
	and every minimal nonzero ideal has the form $M_{n_i}(D_i)$.

	Consider the ideal $k\int^l=I$ which is an ideal since $v\int^l=\varepsilon(v)\int^l\in I$. But then consider that if $v'\int^l\in I$, 
	\[v'\int^l v=\varepsilon(v')\int^l v\in k\int^l\]
	so in fact it is a two-sided ideal. Thus $k\int^l\cong M_{n_i}(D_i)$, so 
	\begin{align*}
		(k\int^l)^2&=k\int^l
		\varepsilon(\int^l)\int^l=(\int^l)^2&=a\int^l
	\end{align*}
	for some nonzero $a$. But then $\varepsilon(\int^l)\ne 0$.

	(b) $\Rightarrow$ (a): Now let $e=\frac{1}{\varepsilon(\int^l)}\int^l$. Then $e^2=e$ and $ke=k\int^l$ is a left $V$-module.
	But then $V=Ve\oplus V(1-e)$ so $ke=k\int^l$ is projective, so $\projdim_v k=0$.

	The last part follows by Lorenz-Lorenz -- $\gldim V=\projdim_V k$.
\end{prf}

\subsection{$\Ext$ groups}
We define these in the usual way through projective resolutions:

Let $A$ be an algebra. Then $_A\calM$ is an abelian cateogry, so $\Hom_A(M,N)\in\Vectk$.
\begin{defn}
	A (covariant) functor $F:{_A\calM}\to{_B\calM}$ is called \textbf{left exact} if, for any short exact sequence
	\[0\to L\to M\to N\to 0\]
	we have the exact sequence
	\[0\to F(L)\to F(M)\to F(N).\]

	Similar for right exact. If both, it's just exact.
\end{defn}
\begin{lem}
	For any $W\in{_A\calM}$, the functor $\Hom(W,-)$ is left exact.
\end{lem}
We also have that $\Hom_A(P,-)$ is exact iff $P$ is projective, so this gives us a nice characterization of projective objects in Abelian categories.
Then the category $\Ch({_A\calM})$ is the category of chain complexes in $A$ modules with chain maps.
\begin{lem}
	$\Ch({_A\calM})$ is an Abelian category.
\end{lem}
\begin{lem}
	A functor $F:{_A\calM}\to{_B\calM}$ can be extended to a functor between the appropriate chain complex categories.
\end{lem}

\subsection{Looking Forward}
On Friday, we may see the following result:
\begin{thm}
	Under some mild hypothesis (including finite dimensionality),
	\[\projdim_V k=\gldim V\]
\end{thm}
\end{document}
